---
title: "figure_2a"
author: "Melissa Spear"
date: "14/07/2020"
output: html_document
---


#####Reading in the admixture run file, subsetting the reference pops, extracting admixed samples 
```{r}
#Read in the admixture file estimates file 
HCHS_SOL_admixture_file<-read.table("/Users/melissaspear/Dropbox/HCHS_SOL/HCHS_SOL_clean_set_with_Reich_NAM_CEU_YRI_hg19_FINAL_030718_forwardstrand_PRUNED.3.Q", header =F, sep = ' ')
names(HCHS_SOL_admixture_file) <- c("YRI","CEU","AI")

#Reading in the fam file accompaniment of the .bed file called for the admixture run 
fam_file = read.table("/Users/melissaspear/Dropbox/HCHS_SOL/HCHS_SOL_clean_set_with_Reich_NAM_CEU_YRI_hg19_FINAL_030718_forwardstrand_PRUNED.fam", header=F, sep = ' ')
names(fam_file)<-c("FID","IID","PID","MID","Sex","Pheno")
fam_file_with_IDs <- cbind(fam_file, HCHS_SOL_admixture_file)

#subset ref pops
CEU<-fam_file_with_IDs[1:104,] #subset the CEU ref pop
YRI<-fam_file_with_IDs[105:211,] #subset the YRI ref pop
AI<-fam_file_with_IDs[212:323,] #subset the AI ref pop 

all_ref<-rbind(CEU,YRI,AI)

#all HCHS/SOL with reference pops 
dim(fam_file_with_IDs)
```

##### Reading in the HCHS/SOL phenotype files and merging with the global ancestry estimates 
```{r}
#Read in the HCHS/SOL phenotype file 
HCHS_SOL_pheno<-read.table("/Users/melissaspear/Dropbox/HCHS_SOL/HCHS_SOL_allphenotypes_ALL_samples_new_birthyear.txt", header= T, sep = '\t')

#merge the admixture estimates with the pheno file 
all_with_pheno=merge(fam_file_with_IDs, HCHS_SOL_pheno, by.x =c("FID"), by.y=c("SUBJECT_ID"))
dim(all_with_pheno) #minus the ref
```

#### The following section is for Figure 2 of the paper. 

Subsetting out the admixed 95 samples 
```{r}
#Extracting out only the 95% admixed samples
admixed_with_pheno<-all_with_pheno[which(all_with_pheno$CEU <= 0.95 & all_with_pheno$YRI <=0.95 & all_with_pheno$AI <=0.95),]

#admixed_with_pheno<-merge(admixed_95, HCHS_SOL_pheno, by.x =c("V1"), by.y=c("SUBJECT_ID"))
dim(admixed_with_pheno)
```

#####Subsetting the admixed 95 HCHS/SOL populations.
These are the sample sizes listed in Supplementary Table 1. 
```{r}
#Dominican subset 
Dominican_admixed<-admixed_with_pheno[which(admixed_with_pheno$BKGRD1_C7 == "0"),]
dim(Dominican_admixed)
```

```{r}
#Central American subset
Central_American_admixed<-admixed_with_pheno[which(admixed_with_pheno$BKGRD1_C7 == "1"),]
dim(Central_American_admixed)
```

```{r}
#Cuban subset
Cuban_admixed<-admixed_with_pheno[which(admixed_with_pheno$BKGRD1_C7 == "2"),]
dim(Cuban_admixed)
```

```{r}
#Mexican subset 
Mexican_admixed<-admixed_with_pheno[which(admixed_with_pheno$BKGRD1_C7 == "3"),]
dim(Mexican_admixed)
```

```{r}
#Puerto Rican subset 
Puerto_Rican_admixed<-admixed_with_pheno[which(admixed_with_pheno$BKGRD1_C7 == "4"),]
dim(Puerto_Rican_admixed)
```

```{r}
#South American subset 
South_American_admixed<-admixed_with_pheno[which(admixed_with_pheno$BKGRD1_C7 == "5"),]
dim(South_American_admixed)
```


```{r}
##Plot Figure 2A

#phenotype file of the admixed 95 mexican samples 
d = read.table("/Users/melissaspear/Dropbox/HCHS_SOL/HCHS_SOL_Mexicans_admixed_95_full_pheno_010919.txt",header=T,sep='\t')

N=nrow(d)

fit = lm(d[,9] ~ d[,284]+log(d[,179])) #linear model: AI ancestry vs birth year


#pdf("Amerindigenous_ancestry_ancestry_points+wiskers_dbins_081319_Figure1C.pdf",height=2.25,width=2.375,pointsize=10)
par(mar=c(2.1,2.1,0.1,0.1))
plot(1,1,type='n',xlim=range(d[,284]),ylim=c(0,1),xlab="",ylab="",xaxt='n',yaxt='n')
abline(h=seq(0,1,by=0.1),col=rgb(0,0,0,0.1))
abline(v=seq(1930,2000,by=5),col=rgb(0,0,0,0.1))
axis(side=1,padj=-2,cex.axis=0.7)
axis(side=2,padj=1.5,cex.axis=0.7, at=c("0.0","0.2","0.4","0.6","0.8","1.0"), labels=c("0","20","40","60","80","100"))
mtext(side=1,"Birth year",line=1.1)
mtext(side=2,"Amerindigenous ancestry (%)",line=1.1)
c=col2rgb(pcol[2])/255;
I1 = which.min(d[,284])
I2 = which.max(d[,284])
points(d[,284],d[,9],col=rgb(c[1],c[2],c[3],0.2),pch=16,cex=0.5)
segments(x0=d[I1,284], x1=d[I2,284],y0=fit$fitted.values[I1],y1=fit$fitted.values[I2],col="white",lwd=3)
segments(x0=d[I1,284], x1=d[I2,284],y0=fit$fitted.values[I1],y1=fit$fitted.values[I2],col=pcol[2],lwd=2)
pw = 10 #partition width, in years
part=seq(min(floor(d[,9]/10)*10),max(ceiling(d[,284]/10)*10),by=pw)
for(i in 1:(length(part)-1)){
  q1 = max(part[i],min(d[,284]));
  q2 = min(part[i+1],max(d[,284]));
  dbin = d[which(d[,284]>q1 & d[,284]<=q2),]
  m = mean(dbin[,9],na.rm=T)
  se = 2*sd(dbin[,9],na.rm=T)/sqrt(nrow(dbin))
  segments(x0=q1,x1=q2,y0=m)
  segments(x0=(q1+q2)/2,y0=m-se,y1=m+se)
  cat(i,q1,q2,m,se,"\n",sep="\t")
}
#dev.off()
```
