---
title: "Global ancestry analyses - Figure 2/Table 1"
author: "Melissa Spear"
date: "12/05/2020"
output: html_document
---

The following section is for Figure 2 and Table 1 of the paper.

#####Reading in the admixture run file, subsetting the reference pops, extracting admixed samples 
```{r}
#Read in the admixture file estimates file 
HCHS_SOL_admixture_file<-read.table("/Users/melissaspear/Dropbox/HCHS_SOL/HCHS_SOL_clean_set_with_Reich_NAM_CEU_YRI_hg19_FINAL_030718_forwardstrand_PRUNED.3.Q", header =F, sep = ' ')
names(HCHS_SOL_admixture_file) <- c("YRI","CEU","AI")

#Reading in the fam file accompaniment of the .bed file called for the admixture run 
fam_file = read.table("/Users/melissaspear/Dropbox/HCHS_SOL/HCHS_SOL_clean_set_with_Reich_NAM_CEU_YRI_hg19_FINAL_030718_forwardstrand_PRUNED.fam", header=F, sep = ' ')
names(fam_file)<-c("FID","IID","PID","MID","Sex","Pheno")
fam_file_with_IDs <- cbind(fam_file, HCHS_SOL_admixture_file)

#subset ref pops
CEU<-fam_file_with_IDs[1:104,] #subset the CEU ref pop
YRI<-fam_file_with_IDs[105:211,] #subset the YRI ref pop
AI<-fam_file_with_IDs[212:323,] #subset the AI ref pop 

all_ref<-rbind(CEU,YRI,AI)

#all HCHS/SOL with reference pops 
dim(fam_file_with_IDs)

```

##### Reading in the HCHS/SOL phenotype files and merging with the global ancestry estimates 
```{r}
#Read in the HCHS/SOL phenotype file 
HCHS_SOL_pheno<-read.table("/Users/melissaspear/Dropbox/HCHS_SOL/HCHS_SOL_allphenotypes_ALL_samples_new_birthyear.txt", header= T, sep = '\t')

#merge the admixture estimates with the pheno file 
all_with_pheno=merge(fam_file_with_IDs, HCHS_SOL_pheno, by.x =c("FID"), by.y=c("SUBJECT_ID"))
dim(all_with_pheno) #minus the ref
```


Subsetting out the admixed 95 samples 
```{r}
#Extracting out only the 95% admixed samples
admixed_with_pheno<-all_with_pheno[which(all_with_pheno$CEU <= 0.95 & all_with_pheno$YRI <=0.95 & all_with_pheno$AI <=0.95),]

#admixed_with_pheno<-merge(admixed_95, HCHS_SOL_pheno, by.x =c("V1"), by.y=c("SUBJECT_ID"))
dim(admixed_with_pheno)
```

#####Subsetting the admixed 95 HCHS/SOL populations.
These are the sample sizes listed in Supplementary Table 1. 
```{r}
#Dominican subset 
Dominican_admixed<-admixed_with_pheno[which(admixed_with_pheno$BKGRD1_C7 == "0"),]
dim(Dominican_admixed)
```

```{r}
#Central American subset
Central_American_admixed<-admixed_with_pheno[which(admixed_with_pheno$BKGRD1_C7 == "1"),]
dim(Central_American_admixed)
```

```{r}
#Cuban subset
Cuban_admixed<-admixed_with_pheno[which(admixed_with_pheno$BKGRD1_C7 == "2"),]
dim(Cuban_admixed)
```

```{r}
#Mexican subset 
Mexican_admixed<-admixed_with_pheno[which(admixed_with_pheno$BKGRD1_C7 == "3"),]
dim(Mexican_admixed)
```

```{r}
#Puerto Rican subset 
Puerto_Rican_admixed<-admixed_with_pheno[which(admixed_with_pheno$BKGRD1_C7 == "4"),]
dim(Puerto_Rican_admixed)
```

```{r}
#South American subset 
South_American_admixed<-admixed_with_pheno[which(admixed_with_pheno$BKGRD1_C7 == "5"),]
dim(South_American_admixed)
```


```{r}
##Plot Figure 2A

#phenotype file of the admixed 95 mexican samples 
d = read.table("/Users/melissaspear/Dropbox/HCHS_SOL/HCHS_SOL_Mexicans_admixed_95_full_pheno_010919.txt",header=T,sep='\t')

N=nrow(d)

fit = lm(d[,9] ~ d[,284]+log(d[,179])) #linear model: AI ancestry vs birth year

#pdf("Amerindigenous_ancestry_ancestry_points+wiskers_dbins_081319_Figure1C.pdf",height=2.25,width=2.375,pointsize=10)
par(mar=c(2.1,2.1,0.1,0.1))
plot(1,1,type='n',xlim=range(d[,284]),ylim=c(0,1),xlab="",ylab="",xaxt='n',yaxt='n')
abline(h=seq(0,1,by=0.1),col=rgb(0,0,0,0.1))
abline(v=seq(1930,2000,by=5),col=rgb(0,0,0,0.1))
axis(side=1,padj=-2,cex.axis=0.7)
axis(side=2,padj=1.5,cex.axis=0.7, at=c("0.0","0.2","0.4","0.6","0.8","1.0"), labels=c("0","20","40","60","80","100"))
mtext(side=1,"Birth year",line=1.1)
mtext(side=2,"Amerindigenous ancestry (%)",line=1.1)
c=col2rgb(pcol[2])/255;
I1 = which.min(d[,284])
I2 = which.max(d[,284])
points(d[,284],d[,9],col=rgb(c[1],c[2],c[3],0.2),pch=16,cex=0.5)
segments(x0=d[I1,284], x1=d[I2,284],y0=fit$fitted.values[I1],y1=fit$fitted.values[I2],col="white",lwd=3)
segments(x0=d[I1,284], x1=d[I2,284],y0=fit$fitted.values[I1],y1=fit$fitted.values[I2],col=pcol[2],lwd=2)
pw = 10 #partition width, in years
part=seq(min(floor(d[,9]/10)*10),max(ceiling(d[,284]/10)*10),by=pw)
for(i in 1:(length(part)-1)){
  q1 = max(part[i],min(d[,284]));
  q2 = min(part[i+1],max(d[,284]));
  dbin = d[which(d[,284]>q1 & d[,284]<=q2),]
  m = mean(dbin[,9],na.rm=T)
  se = 2*sd(dbin[,9],na.rm=T)/sqrt(nrow(dbin))
  segments(x0=q1,x1=q2,y0=m)
  segments(x0=(q1+q2)/2,y0=m-se,y1=m+se)
  cat(i,q1,q2,m,se,"\n",sep="\t")
}
#dev.off()
```


```{r}
#Figure 2B
MX_pheno<-read.table("/Users/melissaspear/Dropbox/HCHS_SOL/HCHS_SOL_Mexicans_admixed_95_full_pheno_010919.txt",header=T,sep='\t')

set.seed(123)
fit.native = lm(NAM ~ estimated_birth_year+ log(WEIGHT_FINAL_NORM_OVERALL),data=MX_pheno)

#Bootstrap iterations of global ancestry 
library(boot)

boot_fn <- function(data, indices) {
  d <- data[indices, ]
  d <- d[order(d$estimated_birth_year),]
  loess_fit <- loess(NAM ~ estimated_birth_year, d,
                     control = loess.control(surface = "direct"))
  predict(loess_fit, data.frame(estimated_birth_year = seq(1934, 1993, 1)), se = T)$fit
}

loess_boot <- boot(MX_pheno, R = 1000, statistic = boot_fn)

``` 

```{r}
pdf("/Users/melissaspear/Dropbox/HCHS_SOL/paper_figures/NativeAmerican_ancestry_bootstrap_loess_linear_081319_new_figure_2B.pdf",height=2.25,width=2.375,pointsize=10)
par(mar=c(2.1,2.1,0.1,0.1))
plot(1,1,type='n',xlim=range(d[,284]),ylim=c(0.38,0.55),xlab="",ylab="",xaxt='n',yaxt='n')
abline(h=seq(0.35,0.6,by=0.05),col=rgb(0,0,0,0.1))
abline(v=seq(1930,2000,by=5),col=rgb(0,0,0,0.1))
axis(side=1,padj=-2,cex.axis=0.60,cex.lab=0.90)
axis(side=2,padj=1.5,cex.axis=0.60,cex.lab=0.90,at=c("0.35","0.4","0.45","0.5","0.55","0.6"), labels=c("35","40","45","50","55","60"))
mtext(side=1,"Birth year",line=1.1)
mtext(side=2,"Amerindigenous ancestry (%)",line=1.1)
#plot(MX_pheno$estimated_birth_year, MX_pheno$NAM,ylim=c(0.3,0.6),pch = 20, col = "black", xlab = "Birth year",
     #ylab = "Native American ancestry (%)")
for(i in sample(nrow(loess_boot$t), 1000)) { #function to plot each row on top of one another 
  lines(seq(1934, 1993, 1), loess_boot$t[i,], col = "gray",lwd=0.75)
}
I1 = which.min(d[,284])
I2 = which.max(d[,284])
segments(x0=d[I1,284], x1=d[I2,284],y0=fit.native$fitted.values[I1],y1=fit.native$fitted.values[I2],col="white",lwd=4)
segments(x0=d[I1,284], x1=d[I2,284],y0=fit.native$fitted.values[I1],y1=fit.native$fitted.values[I2],col=pcol[2],lwd=3)
pw = 10 #partition width, in years
part=seq(min(floor(d[,9]/10)*10),max(ceiling(d[,284]/10)*10),by=pw)
for(i in 1:(length(part)-1)){
  q1 = max(part[i],min(d[,284]));
  q2 = min(part[i+1],max(d[,284]));
  dbin = d[which(d[,284]>q1 & d[,284]<=q2),]
  m = mean(dbin[,9],na.rm=T)
  se = 2*sd(dbin[,9],na.rm=T)/sqrt(nrow(dbin))
  segments(x0=q1,x1=q2,y0=m)
  segments(x0=(q1+q2)/2,y0=m-se,y1=m+se)
  cat(i,q1,q2,m,se,"\n",sep="\t")
}
conf_97.5 <- apply(loess_boot$t, 2, function(x) quantile(x, .975))
conf_2.5 <- apply(loess_boot$t, 2, function(x) quantile(x, .025))
lines(seq(1934, 1993, 1), conf_97.5, type ="l",
      col = "black", lty=5)
lines(seq(1934, 1993, 1), conf_2.5, type ="l",
      col = "black",lty=5)
dev.off()

```



The following section is for Table 1 of the paper 

```{r}
Mexican_reclassified<-read.table("/Users/melissaspear/Dropbox/HCHS_SOL/HCHS_SOL_Mexicans_admixed_95_full_pheno_010919.txt",header=T,sep='\t')

#Migrant status
US_born = Mexican_reclassified[which(Mexican_reclassified$US_BORN == '1'),]
US_migrant = Mexican_reclassified[which(Mexican_reclassified$US_BORN == '0'),]

US_born_model<-lm(NAM ~ estimated_birth_year+log(WEIGHT_FINAL_NORM_OVERALL), data = US_born)
US_migrant_model<-lm(NAM ~ estimated_birth_year+log(WEIGHT_FINAL_NORM_OVERALL), data = US_migrant)

#Region 
Chicago = Mexican_reclassified[which(Mexican_reclassified$CENTER == 'C'),]
San_Diego = Mexican_reclassified[which(Mexican_reclassified$CENTER == 'S'),]
Miami = Mexican_reclassified[which(Mexican_reclassified$CENTER == 'M'),]
Bronx = Mexican_reclassified[which(Mexican_reclassified$CENTER == 'B'),]

Chicago_model<-lm(NAM ~ estimated_birth_year+log(WEIGHT_FINAL_NORM_OVERALL), data = Chicago)
San_Diego_model<-lm(NAM ~ estimated_birth_year+log(WEIGHT_FINAL_NORM_OVERALL), data = San_Diego)

#Educational_attainment
Mexican_less_high_school = Mexican_reclassified[which(Mexican_reclassified$EDUCATION_C3 == '1'),]
Mexican_high_school = Mexican_reclassified[which(Mexican_reclassified$EDUCATION_C3 == '2'),]
Mexican_higher_high_school = Mexican_reclassified[which(Mexican_reclassified$EDUCATION_C3 == '3'),]


Mexican_less_high_school_model<-lm(NAM ~ estimated_birth_year+log(WEIGHT_FINAL_NORM_OVERALL), data = Mexican_less_high_school)
Mexican_high_school_model<-lm(NAM ~ estimated_birth_year+log(WEIGHT_FINAL_NORM_OVERALL), data = Mexican_high_school)
Mexican_higher_high_school_model<-lm(NAM ~ estimated_birth_year+log(WEIGHT_FINAL_NORM_OVERALL), data = Mexican_higher_high_school)

#Sex
male = Mexican_reclassified[which(Mexican_reclassified$GENDER == 'M'),]
female = Mexican_reclassified[which(Mexican_reclassified$GENDER == 'F'),]

male_model<-lm(NAM ~ estimated_birth_year+log(WEIGHT_FINAL_NORM_OVERALL), data = male)
female_model<-lm(NAM ~ estimated_birth_year+log(WEIGHT_FINAL_NORM_OVERALL), data = female)


###
All<-setNames(data.frame(matrix(ncol = 8, nrow = 0)), c("Category", "N", "Mean","Median","R2","Effect","Std.Err","P"))
 
All[1,1]<-"All"
All[1,2]<-dim(d)[1]
All[1,3]<-mean(d$NAM)
All[1,4]<-median(d$NAM)
All[1,5]<-summary(fit)$r.squared
All[1,6]<-summary(fit)$coeff[2,1]
All[1,7]<-summary(fit)$coeff[2,2]
All[1,8]<-summary(fit)$coeff[2,4]

#chicago
Chicago_df<-setNames(data.frame(matrix(ncol = 8, nrow = 0)), c("Category", "N", "Mean","Median","R2","Effect","Std.Err","P"))

Chicago_df[1,1]<-"Chicago"
Chicago_df[1,2]<-dim(Chicago)[1]
Chicago_df[1,3]<-mean(Chicago$NAM)
Chicago_df[1,4]<-median(Chicago$NAM)
Chicago_df[1,5]<-summary(Chicago_model)$r.squared
Chicago_df[1,6]<-summary(Chicago_model)$coeff[2,1]
Chicago_df[1,7]<-summary(Chicago_model)$coeff[2,2]
Chicago_df[1,8]<-summary(Chicago_model)$coeff[2,4]

#San diego
San_Diego_df<-setNames(data.frame(matrix(ncol = 8, nrow = 0)), c("Category", "N", "Mean","Median","R2","Effect","Std.Err","P"))

San_Diego_df[1,1]<-"San Diego"
San_Diego_df[1,2]<-dim(San_Diego)[1]
San_Diego_df[1,3]<-mean(San_Diego$NAM)
San_Diego_df[1,4]<-median(San_Diego$NAM)
San_Diego_df[1,5]<-summary(San_Diego_model)$r.squared
San_Diego_df[1,6]<-summary(San_Diego_model)$coeff[2,1]
San_Diego_df[1,7]<-summary(San_Diego_model)$coeff[2,2]
San_Diego_df[1,8]<-summary(San_Diego_model)$coeff[2,4]

#US born 
US_born_df<-setNames(data.frame(matrix(ncol = 8, nrow = 0)), c("Category", "N", "Mean","Median","R2","Effect","Std.Err","P"))

US_born_df[1,1]<-"US born"
US_born_df[1,2]<-dim(US_born)[1]
US_born_df[1,3]<-mean(US_born$NAM)
US_born_df[1,4]<-median(US_born$NAM)
US_born_df[1,5]<-summary(US_born_model)$r.squared
US_born_df[1,6]<-summary(US_born_model)$coeff[2,1]
US_born_df[1,7]<-summary(US_born_model)$coeff[2,2]
US_born_df[1,8]<-summary(US_born_model)$coeff[2,4]

#US Migrant
US_migrant_df<-setNames(data.frame(matrix(ncol = 8, nrow = 0)), c("Category", "N", "Mean","Median","R2","Effect","Std.Err","P"))

US_migrant_df[1,1]<-"US migrant"
US_migrant_df[1,2]<-dim(US_migrant)[1]
US_migrant_df[1,3]<-mean(US_migrant$NAM)
US_migrant_df[1,4]<-median(US_migrant$NAM)
US_migrant_df[1,5]<-summary(US_migrant_model)$r.squared
US_migrant_df[1,6]<-summary(US_migrant_model)$coeff[2,1]
US_migrant_df[1,7]<-summary(US_migrant_model)$coeff[2,2]
US_migrant_df[1,8]<-summary(US_migrant_model)$coeff[2,4]

#Male 

male_df<-setNames(data.frame(matrix(ncol = 8, nrow = 0)), c("Category", "N", "Mean","Median","R2","Effect","Std.Err","P"))

male_df[1,1]<-"Male"
male_df[1,2]<-dim(male)[1]
male_df[1,3]<-mean(male$NAM)
male_df[1,4]<-median(male$NAM)
male_df[1,5]<-summary(male_model)$r.squared
male_df[1,6]<-summary(male_model)$coeff[2,1]
male_df[1,7]<-summary(male_model)$coeff[2,2]
male_df[1,8]<-summary(male_model)$coeff[2,4]

#Female
female_df<-setNames(data.frame(matrix(ncol = 8, nrow = 0)), c("Category", "N", "Mean","Median","R2","Effect","Std.Err","P"))

female_df[1,1]<-"Female"
female_df[1,2]<-dim(female)[1]
female_df[1,3]<-mean(female$NAM)
female_df[1,4]<-median(female$NAM)
female_df[1,5]<-summary(female_model)$r.squared
female_df[1,6]<-summary(female_model)$coeff[2,1]
female_df[1,7]<-summary(female_model)$coeff[2,2]
female_df[1,8]<-summary(female_model)$coeff[2,4]

#less than high school
Mexican_less_high_school_df<-setNames(data.frame(matrix(ncol = 8, nrow = 0)), c("Category", "N", "Mean","Median","R2","Effect","Std.Err","P"))

Mexican_less_high_school_df[1,1]<-"<HS"
Mexican_less_high_school_df[1,2]<-dim(Mexican_less_high_school)[1]
Mexican_less_high_school_df[1,3]<-mean(Mexican_less_high_school$NAM)
Mexican_less_high_school_df[1,4]<-median(Mexican_less_high_school$NAM)
Mexican_less_high_school_df[1,5]<-summary(Mexican_less_high_school_model)$r.squared
Mexican_less_high_school_df[1,6]<-summary(Mexican_less_high_school_model)$coeff[2,1]
Mexican_less_high_school_df[1,7]<-summary(Mexican_less_high_school_model)$coeff[2,2]
Mexican_less_high_school_df[1,8]<-summary(Mexican_less_high_school_model)$coeff[2,4]

#high school equiv
Mexican_high_school_df<-setNames(data.frame(matrix(ncol = 8, nrow = 0)), c("Category", "N", "Mean","Median","R2","Effect","Std.Err","P"))

Mexican_high_school_df[1,1]<-"=HS"
Mexican_high_school_df[1,2]<-dim(Mexican_high_school)[1]
Mexican_high_school_df[1,3]<-mean(Mexican_high_school$NAM)
Mexican_high_school_df[1,4]<-median(Mexican_high_school$NAM)
Mexican_high_school_df[1,5]<-summary(Mexican_high_school_model)$r.squared
Mexican_high_school_df[1,6]<-summary(Mexican_high_school_model)$coeff[2,1]
Mexican_high_school_df[1,7]<-summary(Mexican_high_school_model)$coeff[2,2]
Mexican_high_school_df[1,8]<-summary(Mexican_high_school_model)$coeff[2,4]

#Greater than high school 
Mexican_higher_high_school_df<-setNames(data.frame(matrix(ncol = 8, nrow = 0)), c("Category", "N", "Mean","Median","R2","Effect","Std.Err","P"))

Mexican_higher_high_school_df[1,1]<-">HS"
Mexican_higher_high_school_df[1,2]<-dim(Mexican_higher_high_school)[1]
Mexican_higher_high_school_df[1,3]<-mean(Mexican_higher_high_school$NAM)
Mexican_higher_high_school_df[1,4]<-median(Mexican_higher_high_school$NAM)
Mexican_higher_high_school_df[1,5]<-summary(Mexican_higher_high_school_model)$r.squared
Mexican_higher_high_school_df[1,6]<-summary(Mexican_higher_high_school_model)$coeff[2,1]
Mexican_higher_high_school_df[1,7]<-summary(Mexican_higher_high_school_model)$coeff[2,2]
Mexican_higher_high_school_df[1,8]<-summary(Mexican_higher_high_school_model)$coeff[2,4]


###Merging all models 
all_models<-rbind(All,Chicago_df,San_Diego_df,US_born_df,US_migrant_df,male_df,female_df,Mexican_less_high_school_df,Mexican_high_school_df,Mexican_higher_high_school_df)

all_models

```


